Palomar 5 Gaps

### 2024-09-16 paper/main.tex
trying to solve problems with noise

\documentclass[]{aa}

@@

    Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster. The y'coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.

    \subsubsection*{Density maps and profiles}
      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system. This is done by creating 2D density maps of the streams, followed by marginalizing over the y'-coordinate to produce 1D density profiles along the x'-axis. These comparisons are illustrated in the results section (see Fig.~XXX).

      To construct the 1D density profiles, we first bin the data using the  $\sqrt{N}$-rule for determining the number of bins. After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs. This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~XX, where the center of mass has been effectively removed.

      With the smoothed 1D profiles in hand, we proceed to compare them by identifying regions where the full simulations (GCGC) show significant underdensity compared to the \texttt{vanilla} simulations. To achieve this, we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean. However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50. This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters. The identified gap candidates are marked and labeled on the plots.

>>>>

\documentclass[draft]{aa}

@@

    Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster. The y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.

    \subsubsection*{Density maps and profiles}
      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system. We inspect these differences by creating 2D density maps of the streams. Subsequently, we wish to employ a quantitative method to make detections discover when the \textit{full} simulations are significantly less dense than the \textit{vanilla} simulations. To do so, we marginalize over the y'-coordinate to produce 1D density profiles along the x'-axis. These comparisons are illustrated in the results section (see Fig.~\ref{fig:profiles}).

      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$. After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs. This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.

      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance. To do so, we first compute the log ratio of the two profiles:
      \begin{equation}
        \mathcal{R} = \log\left(\frac{\rho_f}{\rho_v}\right),
        \end{equation}
      where $\rho$ is the 1D density, $v$ corresponding to the vanilla simulation, and $f$ corresponding to the full simulation, and $R$ is the log ratio of the two. Using the log ratio is important, because the distribution is symmetric and zero centered. For example if we only used the ratio, then the codomain would be $\frac{\rho_f}{\rho_v} \in [1,+\inf]$ for $\rho_f > \rho_v$, and $\frac{\rho_f}{\rho_v} \in [0,1]$ for $\rho_f < \rho_v$.

      Yet, we only wish to consider bins with a significant amount of counts. We find this threshold by asserting that the signal to noise ratio must be greater than 50. The signal is based on the \textit{vanilla} profile, and the noise is the error propogation given poisson uncertainties for the number of counts a given bin: 
      \begin{equation}
        \mathcal{S}/\mathcal{N} =50= \frac{\log{N}}{\frac{1}{\ln(10)} \sqrt{N}}
        \end{equation}

      we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean. However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50. This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters. The identified gap candidates are marked and labeled on the plots.
